    ---
    # title: "Reproducible Research: Peer Assessment1"
    author: "Chiagozie-Umeano"
    date: "10/21/2020"
    output: html_document
    keep_md: TRUE
    ---

    ```{r setup, include=TRUE}
    knitr::opts_chunk$set(echo = TRUE)
    knitr::opts_chunk$set(fig.path = "./figure/")
    knitr::opts_chunk$set(fig.keep = "all")
    ```

   ## Loading and preprocessing the data
   ======================================
   Set the working directory, the dataset is already in the forked GitHub repository, unzip the file.

   ```{r load data}
   unzip("./activity.zip")
   activityData<-read.csv("./activity.csv")
   summary(activityData)
   ```

    The variable names and entire file structure
    
    ```{r activity details}
    names(activityData)
    head(activityData)
    str(activityData)
    ```
    
    The file is now ready for analysis.

    ## Question 1: What is the mean total number of steps taken per day?
    Ignoring the missing values in the dataset

    1. The total number of steps taken per day
    
    ```{r total steps}
    stepsPerDay<-aggregate(steps~date, activityData, sum, na.rm=TRUE)
    ```

    2. Histogram of total steps per day
    
    ```{r histogram of total steps per day}
    png("plot1.png")
    hist(stepsPerDay$steps)
    dev.off()
    ```

    3. Mean and median of total number of steps taken per day
    
    ```{r mean and median of total steps per day}
    meanStepsPerDay<-mean(stepsPerDay$steps)
    meanStepsPerDay
    medianStepsPerDay<-median(stepsPerDay$steps)
    medianStepsPerDay
    ```

    ## Question 2: What is the average daily activity pattern?

    1. Time series plot (type= "1") of the 5-minute interval(x-axis)  and average number of steps taken, averaged across all days(y-axis)
    
    ```{r time series plot}
    png("plot2.png")
    stepsPerInterval<-aggregate(steps~interval, data=activityData, mean, na.rm=TRUE)
    plot(steps~interval, data=stepsPerInterval, type="l")
    dev.off()
    ```

    2. 5-minute interval on average across all the days in the dataset     contains the maximum number of steps
    
    ```{r 5-minute interval with the maximum number of steps}
    intervalWithMaxNbSteps <- stepsPerInterval[which.max(stepsPerInterval$steps),]$interval
    intervalWithMaxNbSteps
    ```

    ## Question 3: Imputing missing values

    1. Total number of missing values in the dataset
    
    ```{r total number of rows with NAs}
    totalValuesMissings <- sum(is.na(activityData$steps))
    totalValuesMissings
    ```

    2. Strategy for filling in all missing values in the dataset
    
    ```{r filling in all missing values using mean per interval}
    getMeanStepsPerInterval<-function(interval){
    stepsPerInterval[stepsPerInterval$interval==interval,]$steps
    }
    ```

    3. Creating a new dataset that is equal to the original dataset but with the missing data filled in
    
    ```{r new dataset with filled in missing values}
    activityDataNoNA<-activityData
    for(i in 1:nrow(activityDataNoNA)){
    if(is.na(activityDataNoNA[i,]$steps)){
      activityDataNoNA[i,]$steps <- getMeanStepsPerInterval(activityDataNoNA[i,]$interval)
    }
    }
    ```

    4. Histogram of the total number of steps taken, mean and median      of total number of steps taken per day
    
    ```{r histogram of total number of steps taken, mean and median}
    png("plot3.png")
    totalStepsPerDayNoNA <- aggregate(steps ~ date, data=activityDataNoNA, sum)
    hist(totalStepsPerDayNoNA$steps)
    dev.off()
    meanStepsPerDayNoNA <- mean(totalStepsPerDayNoNA$steps)
    medianStepsPerDayNoNA <- median(totalStepsPerDayNoNA$steps)
    ```

    Do the values differ from values in the first assignment? 
    
    The mean of the total steps per day and the mean of filled missing data(NAs) are the same (10766.1886792453). So no change.
    But
    The median of the total steps per day and the median of filled   missing data(NAs) are slightly different. There is a slight shift (from  10765 to 10766.1886792453). A change of 0.1%.
 
    What is the impact of imputing missing data on the estimates of the total daily number of steps?
    
    The impact was only seen on the median, and was negligible. So missing values can be ignored in analysis without much impact in  computation.
 
    ## Question 4: Are there differences in  activity patterns between weekdays and weekends?
    
    Using the weekdays()function and dataset with the filled-in missing values.
 
    1. Creating a new factor variable in the dataset with two levels "weekday" and "weekend".
    
    ```{r creating new factor variable-"weekday" and "weekend"}
    activityDataNoNA$date <- as.Date(strptime(activityDataNoNA$date, format="%Y-%m-%d"))
    activityDataNoNA$day <- weekdays(activityDataNoNA$date)
  for (i in 1:nrow(activityDataNoNA)) {
    if (activityDataNoNA[i,]$day %in% c("Saturday","Sunday")) {
        activityDataNoNA[i,]$day<-"weekend"}
    else{
        activityDataNoNA[i,]$day<-"weekday"}}
    stepsByDay <- aggregate(activityDataNoNA$steps ~  activityDataNoNA$interval + activityDataNoNA$day, activityDataNoNA, mean)
     ```

   2. A panel plot containing a time series plot (type "1") of 5-minute interval(x-axis) and the average number of steps taken.
     
   ```{r panel plot}
   png("plot4.png")
   names(stepsByDay) <- c("interval", "day", "steps")
   library(lattice)
   xyplot(steps ~ interval | day, stepsByDay, type = "l", layout =  c(1, 2), 
    xlab = "Interval", ylab = "Number of steps")
  dev.off()
  ```
}